---
title: "BqExAvI Survey Analysis (2017-2022)"
author: "Modesto"
date: "`r Sys.Date()`"
format:
  html:
    page-layout: full
    toc: true
    toc-location: left
    toc-depth: 2
    number-sections: true
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
    link-external-icon: true
    link-external-newwindow: true 
#output:
#  pdf_document:
#    latex_engine: xelatex
---

# Contents and Disclaimer

```{r}

```

This file contains the data from the "Experimental Advanced Biochemistry I" course (Biochemistry Degree, [Universidad Autónoma de Madrid](https://www.uam.es)). Over the last years, a number of lecturers and professors from the [Biochemistry Department](https://www.bq.uam.es/Index.php) have participated in this course, as follows (in alphabetical order): Julián Aragonés, Juan J. Arredondo, Víctor Calvo, José G. Castaño, Alicia González-Martín, Benilde Jiménez, Marina Lasa, Óscar Martínez-Costa, Luis del Peso, Modesto Redrejo-Rodríguez, Ana I. Rojo, Alejandro Samhan-Arias, and Isabel Sánchez-Pérez.

The questionnaires were completed by the students in 2017-2022 directly on Moodle on the last day of each course. This is a preliminary summary of the data analysis. The GitHub [repo](https://github.com/mredrejo/bqexav) contains the original files of all analyzes. This report only contains the data analysis and plots, without any results discussion. The companion manuscript will be available soon.

The data is made available under the Creative Common License (CC BY-NC-ND 3.0 ES). 

::: {.callout-warning}
## Preliminary
This is only a preliminary analysis. Contact [modesto.redrejo\@uam.es](mailto:modesto.redrejo@uam.es) or [juan.arredondo\@uam.es](mailto:juan.arrredondo@uam.es) for any feedback or queries.
:::



# Moodle survey

```{r results='hide', message=FALSE, warning=FALSE}
#Load/install requires packages
paquetes <- c("ggplot2","data.table","kableExtra","corrplot","likert","ggpubr","heatmaply","reshape2","plotly","dplyr")
unavailable <- setdiff(paquetes, rownames(installed.packages()))
invisible(install.packages(unavailable))
invisible(lapply(paquetes, library, character.only = TRUE))
```

We import data from student responses in 5 years, between 2018-2022. The quiz consists of up to 77 questions, including 5 free text questions (50, 51, 52, 75, 76 & 77), 7 questions with three options, and the rest as a 5-degrees *Likert* scale.

```{r warning=FALSE}
#load questions
questions <- read.csv("questions.csv", head=TRUE, sep=";")
questions <- cbind(row.names(questions),questions[,c(3,1,2)])
#add type variable
questions$type <- "Pos."
questions$type[questions$Section=="Open"] <- ""
questions[c(4,10,56,63,63,64,65,66,67,72),5] <- "Neg."
questions$type[questions$Section=="Open"] <- "NA"
colnames(questions) <- c("No.","Since","Section","Question","Type")
#write.csv(questions,"questions_final.csv", row.names=FALSE)

#diplay the table
kbl(questions[,1:4], align = "cccl", caption = "Table 1. Students opinion quizz. The bulk of the questionaire was designed for the year 2017 and new questions were added as indicated.") %>%
    kable_styling(bootstrap_options = "striped", full_width = F) %>%
    column_spec(1, italic = T)
```

# Load data and pairwise t.test

Survey responses were downloaded from Moodle as txt/csv files. Moodle updates caused some format differences that could be worked around after opening the files with Numbers and exporting them as tables with ";" as the column separator. We now show all vs. all pairwise t-tests performed to detect significant differences in responses per year. The tables below contain the pairwise p-values for each comparison, with significant values highlighted in blue (p\<0.05) and red (p\<0.01)

```{r warning=FALSE, results='asis'}

#read the data in a list of dataframes
#didn't use the headers to avoid mistakes
quiz <- lapply(2017:2022, function(x) read.csv(paste0("survey",x,".csv"),header=FALSE,skip=1,sep=";"))
#add Year as the third variable (empty so far)
curso <- c("2017","2018","2019","2020","2021","2022")
for (i in 1:length(quiz)){
  quiz[[i]][,3] <- curso[i]
}

#adjust questions changes
#remove questions in column 32 & 40 from 2017, because we removed it in the following years
quiz[[1]] <- quiz[[1]][,-c(32,40)]
quiz[[6]] <- quiz[[6]][,-10]
names(quiz[[6]]) <- names(quiz[[5]])
names(quiz[[1]]) <- names(quiz[[2]])
#merge all dataframes and name the columns 
data <- Reduce(function(x, y) merge(x, y, all=TRUE), quiz)
#take the colnames from the last quiz that contains all the questions

colnames(data)[3] <- "Curso"


#statistics analysis
#subset questions 1/2: remove leftmost junk columns
subdata_all <- data[,c(3,11:87)]
names(subdata_all) <- c("Curso",paste0("Q",1:77))
#write.csv(subdata_all, "merged_data.csv")

#subset questions 2/2: remove open questions
open <- c(row.names(questions[questions$Section=="Open",]))
subdata <- subdata_all[,-(as.integer(open)+1)]
subdata <- sapply(subdata,as.numeric)
subdata <- as.data.frame(subdata)

#perform tests and display table in a loop
tests <- list()
nombres <- c()

for (i in 2:ncol(subdata)){
  subdata[,i][!(subdata[,i] %in% c(1,2,3,4,5))] <- NA
  #subset for years with answers to avoid void groups
  kkk <- subset(subdata,!is.na(subdata[,i]))
  tests[[i-1]] <- pairwise.t.test(x=as.numeric(kkk[,i]),g=as.numeric(kkk[,1]),paired = F)
  nombres[i-1] <- questions[,4][(which(questions$No. %in% gsub("\\D", "",colnames(subdata[i]))))]
  names(tests) <- nombres
  print(as.data.frame(format(tests[[i-1]]$p.value, scientific=F,nsmall=6)) %>% replace(., . < 0, "") %>%
    mutate_all(~cell_spec(.x, color = ifelse(.x < 0.01, "firebrick", ifelse(.x < 0.05, "steelblue",
        "black")))) %>%
    kable(escape = F, align = "cccl", caption =paste("<b>",names(tests[i-1]),"</b>"), digits=4)  %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "center") %>%
    column_spec(1, bold = T))
  
  
  
}

```

# Likert scale plots

*Likert* scale responses are grouped by the sections in the quiz.

## General Methodology

```{r Fig2, echo=TRUE, fig.height=7,warning=FALSE}

#lickert
#change question names
tablita <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = length(colnames(subdata)),
                          ncol = 2))
for (i in 2:length(colnames(subdata))){
  tablita[i-1,] <- cbind(colnames(subdata[i]),questions$Question[as.numeric(gsub("\\D", "",colnames(subdata[i])))])
  colnames(subdata)[i] <- tablita[i-1,2]
}

for (i in 2:72){
  subdata[,i] <- factor(subdata[,i])
}
subdata$Curso <- factor(subdata$Curso,levels=c(2022,2021,2020,2019,2018,2017))
#questions with 3 options
#General Methodology
subdata[,c(8:12)] <- lapply(subdata[,c(8:12)], function(x) factor(x, 
      labels = c("Traditional","Same","Open Question"))
  )
xlikgroup3a = likert(subdata[,c(8:12)], grouping = subdata$Curso)
plot(xlikgroup3a, type = "bar", centered = T) 
```

## Activities Length

```{r Fig3, echo=TRUE, fig.height=10,warning=FALSE}
subdata[,c(29:35)] <- lapply(subdata[,c(29:35)], function(x) factor(x, 
      labels = c("Less","Fine","More"))
  )

xlikgroup3b = likert(subdata[,c(29:35)], grouping = subdata$Curso)

plot(xlikgroup3b, type = "bar", centered = T)
#title(main = "Activities Length", xlab = "X axis", ylab = "Y axis", cex.main = 4,   font.main = 3)
#legend("bottom",  c("Less","Fine","More"))
```

## Equipment

```{r Fig4, echo=TRUE,warning=FALSE}
subdata[,c(2:4)] <- lapply(subdata[,c(2:4)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5a = likert(subdata[,c(2:4)], grouping = subdata$Curso)
plot(xlikgroup5a, type = "bar", centered = T)

```

## Length and Schedule

```{r Fig5, echo=TRUE, fig.height=5,warning=FALSE}
subdata[,c(5:7)] <- lapply(subdata[,c(5:7)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
xlikgroup5b = likert(subdata[,c(5:7)], grouping = subdata$Curso)
plot(xlikgroup5b, type = "bar", centered = T)

```

## Method Objectives

```{r Fig6, echo=TRUE, fig.height=13,warning=FALSE}
subdata[,c(13:21)] <- lapply(subdata[,c(13:21)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
xlikgroup5c = likert(subdata[,c(13:21)], grouping = subdata$Curso)
plot(xlikgroup5c, type = "bar", centered = T)

```

## Activities Interest

```{r Fig7, echo=TRUE, fig.height=11, warning=FALSE}
subdata[,c(22:28)] <- lapply(subdata[,c(22:28)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5d = likert(subdata[,c(22:28)], grouping = subdata$Curso)
plot(xlikgroup5d, type = "bar", centered = T)

```

## Assessment

```{r Fig8, echo=TRUE, fig.height=9,warning=FALSE}
subdata[,c(36:41)] <- lapply(subdata[,c(36:41)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5e = likert(subdata[,c(36:41)], grouping = subdata$Curso)
plot(xlikgroup5e, type = "bar", centered = T)

```

## Learning Objectives

```{r Fig9, echo=TRUE, fig.height=12,warning=FALSE}
subdata[,c(42:50)] <- lapply(subdata[,c(42:50)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5f = likert(subdata[,c(42:50)], grouping = subdata$Curso)
plot(xlikgroup5f, type = "bar", centered = T)

```

## ELN

```{r Fig10, echo=TRUE, fig.height=28,warning=FALSE}
#subset to remove empty years
subdata[,c(51:72)] <- lapply(subdata[,c(51:72)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
eln <- subset(subdata[,c(1,51:72)][subdata$Curso==2019|subdata$Curso==2020|subdata$Curso==2021|subdata$Curso==2022,])


xlikgroup5g = likert(eln[2:23], grouping = eln$Curso)
plot(xlikgroup5g, type = "bar", centered = T, title="ELN")
```

# [Open questions](surveys_free_text.html)

Free-text questions have been analyzed independently by automatic text *lemmatization* and plot methods. The detailed report can be found [here](surveys_free_text.html).

# Acknowledgments

This work has been supported by UAM Teaching Innovation Grants (M-015.17-INN, M-020.18-IMP, M_002.19_INN and M_009.20_IMP).

### Session Info

```{r}
sessionInfo()
```
