---
title: "BqExAvI Survey Analysis (2017-2022). Part II"
author: "Modesto"
date: "`r Sys.Date()`"
format:
  html:
    page-layout: full
    toc: true
    toc-location: left
    toc-depth: 2
    number-sections: true
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
#output:
#  pdf_document:
#    latex_engine: xelatex
---

```{r setup, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load R libraries
paquetes <- c("deeplr", "ggplot2", "BTM", "ggraph", "igraph", "dplyr", "akc","tidytext","udpipe","textrank","ggwordcloud","ggpubr")
unavailable <- setdiff(paquetes, rownames(installed.packages()))
install.packages(unavailable)
lapply(paquetes, library, character.only = TRUE,quietly = TRUE)

```

# Contents and Disclaimer
This page refers to student surveys of "Experimental Advanced Biochemistry I" course (Biochemistry Degree, [Universidad Autónoma de Madrid](https://www.uam.es)). Description of questionnaires and analysis of quantitative questions is in the main page [here](https://mredrejo.github.io/bqexav/). The GitHub [repo](https://github.com/mredrejo/bqexav) contains the original files of all analyzes.

The data is made available under the Creative Common License (CC BY-NC-ND 3.0 ES).  

::: {.callout-warning}
## Preliminary
This is only a preliminary analysis. Contact [modesto.redrejo\@uam.es](mailto:modesto.redrejo@uam.es) for more information.
:::

# Free-text questions

The list of questions is available in the companion [report](index.html#moodle-survey). Free-text questions (50, 51, 52, 75, 76, 77) will be processed to identify more common keywords and terms co-ocurrence. First, text are translated with `deeplr` package and save it into a new table.

```{r translate, include=TRUE, eval=FALSE}
#load the data
data <- read.csv("merged_data.csv")
#to deal with the deeplr words limit, each question is translated independently
Q50 <- toEnglish2(data$Q50, auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
Q51 <- toEnglish2(data$Q51, auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
Q52 <- toEnglish2(data$Q52, auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
Q75 <- toEnglish2(na.omit(data$Q75), auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
Q76 <- toEnglish2(na.omit(data$Q76), auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
Q77 <- toEnglish2(na.omit(data$Q77), auth_key = "9c9bf91e-f863-9406-d285-2f209fc1828d:fx")
#subset data and sweep sp to en
data_en <- data[,c(2,52,53,54,77,78,79)]
data_en$Q50 <- Q50
data_en$Q51 <- Q51
data_en$Q52 <- Q52
data_en$Q75[which(!is.na(data_en$Q75))] <- Q75
data_en$Q76[which(!is.na(data_en$Q76))] <- Q76
data_en$Q77[which(!is.na(data_en$Q77))] <- Q77

write.csv(data_en,"survey_open_en.csv")
```

Then, we processed the table and process it with `akc` and `udpipe` R packages. The code is largely based in an `akc` [tutorial](https://hope-data-science.github.io/akc/articles/tutorial_raw_text.html) and `updpipe` [extended vignette](https://cran.r-project.org/web/packages/udpipe/vignettes/udpipe-usecase-postagging-lemmatisation.html).

Previously, but after some rounds of *try and error*, we also performed some *ad hoc* word switching to unify terms (e.g. combine lab and laboratory, remove instructors' names, etc.). We found that the combination of the new `akc` package and `udpipe` achieved the best automatic term extraction and lemmatization with improved selection of key terms with reduced noise (conjunctions and other extra words). However, as `akc` is designed for text network-based classification which is beyond our scope, so we decided to implement `udpipe` network plot. Each plot contains the connections of the top-40 most common keywords (nouns and verbs).

```{r i, fig.height=40,fig.width=8, warning=FALSE, message=FALSE}

q_en <-  read.csv("survey_open_en.csv",row.names = NULL, stringsAsFactors = FALSE)

#naif combination of some identical terms
q_en  <- as.data.frame(sapply(q_en,tolower))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('&nbsp;','',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('laboratory','lab',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('script','protocol',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('team','group',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('groupwork','group',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('learning','learn',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('sandra','assistants',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('barbara','assistants',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('barbara acosta','assistants',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('rojo','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub(' ana ', 'instructors',x)))
#since "ana" is within some words, I also changed "ANA" and "Ana Rojo" to rojo manually
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('luis','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('juan','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('modesto','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('benilde','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub("\\\\","",x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('oscar','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('teachers','instructors',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('internship','practices',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('internshipss','practices',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('script','protocol',x)))
q_en <- data.frame(apply(q_en, 2, function(x)  gsub('notebooks','notebook',x)))

#clean and extrack keywords
#from https://hope-data-science.github.io/akc/articles/tutorial_raw_text.html
# https://cran.r-project.org/web/packages/akc/vignettes/akc_vignette.html

#model with udpipe & create dictionary   

#download the model the first time
#ud_model <- udpipe_download_model(language = "english")
#ud_model <- udpipe_load_model(ud_model$file_model)
#load the model from HD
ud_model <- udpipe_load_model("/Users/modesto/Documents/GitHub/bqexav/english-ewt-ud-2.5-191206.udpipe")
x <- list()
Q <- c("Q50","Q51","Q52","Q75","Q76","Q77")
for (i in 1:6){
  x[[i]] <- as.data.frame(udpipe_annotate(ud_model, x = q_en[,i+2], tagger="default",parser="none"))
  names(x)[[i]] <- Q[i]
}


# get stop words from `tidytext` package
stop_words %>%
    pull(word) %>%
    unique() -> my_stopword

#make dictionaries
my_dict <- list()
for (i in 1:6){
  my_dict[[i]] <- make_dict(x[[i]]$lemma)
  #names(my_dict)[[i]] <- Q[i]
}

#extract keywords with akc
extracted_keywords <- list()
for (i in 1:6){
  q_en %>%
    keyword_extract(id = "X",text = colnames(q_en)[i+2],
    dict = my_dict[[i]],stopword = my_stopword) -> extracted_keywords[[i]]
    names(extracted_keywords)[[i]] <- Q[i]
}

#clean data
clean_data <- list()
for (i in 1:6){
  extracted_keywords[[i]] %>%
    keyword_clean() -> clean_data[[i]]
    names(clean_data)[[i]] <- Q[i]
}



#visualize network as in udpipe vignette

#plot from akc extracted_keywords

questions <- read.csv("questions_final.csv", head=TRUE, sep=",") #load questions
titles <- questions[c(50,51,52,75,76,77),4]
stats <- list()
wordnetwork <- list()
plotkey <- list()
for (i in 1:6){
  stats[[i]] <- cooccurrence(x = clean_data[[i]], 
                     term = "keyword", group = "id")
  wordnetwork[[i]] <- head(stats[[i]], 40)
  wordnetwork[[i]] <- graph_from_data_frame(wordnetwork[[i]])
  plotkey[[i]] <- ggraph(wordnetwork[[i]], layout = "fr") +
    geom_edge_link(aes(width = cooc, edge_alpha = cooc), edge_colour = "indianred2") +
    geom_node_text(aes(label = name), col = "darkgreen", size = 7, repel=T,box.padding=0.06) +
    theme_graph(base_family = "Arial", plot_margin = margin(0, 30, 0, 30)) +
    theme(legend.position = "none", plot.margin=unit(c(0,0,2,0), "cm")) +
    labs(title =titles[[i]], subtitle = "Cooccurrences within same sentence (Nouns & Adjective)") 

}
#show the plots
ggarrange(plotlist=plotkey,nrow=6,ncol=1)

```

# Answers per year

Now we split the plots per year. To facilitate the understanding, we have selected the top-20 words.

```{r fig.height=150,fig.width=8, warning=FALSE, message=FALSE}
#extract keywords
extracted_keywords <- list()
indice <- matrix(1:36, byrow = TRUE, nrow = 6)
for (i in 1:6){
  for (j in 1:6){
    tablita <-  q_en[q_en$Curso == 2016+j,]
    if (any(is.na(tablita[,i+2][tablita$Curso == 2016+j]))) next
    tablita %>%
      keyword_extract(id = "X",text = colnames(tablita)[i+2],
      dict = my_dict[[i]], stopword = my_stopword) -> extracted_keywords[[indice[i,j]]]
      names(extracted_keywords)[[indice[i,j]]] <- Q[i]
  }
}


#clean data
clean_data <- list()
for (i in 1:36){
  if (is.null(extracted_keywords[[i]])) next
  extracted_keywords[[i]] %>%
    keyword_clean() -> clean_data[[i]]
    names(clean_data)[[i]] <- Q[i]
}



#visualize

#plot from akc extracted_keywords

year <- rep(unique(q_en$Curso), 6)
stats <- list()
wordnetwork <- list()
plotkey <- list()
for (i in 1:36){
  if (is.null(clean_data[[i]])) next
  stats[[i]] <- cooccurrence(x = clean_data[[i]], 
                     term = "keyword", group = "id")
  wordnetwork[[i]] <- head(stats[[i]], 20)
  wordnetwork[[i]] <- graph_from_data_frame(wordnetwork[[i]])
  plotkey[[i]] <- ggraph(wordnetwork[[i]], layout = "fr") +
    geom_edge_link(aes(width = cooc, edge_alpha = cooc), edge_colour = "indianred2") +
    geom_node_text(aes(label = name), col = "darkgreen", size = 6,repel=T,box.padding=0.06) +
    theme_graph(base_family = "Arial", plot_margin = margin(0, 15, 0, 15)) +
    theme(legend.position = "none", plot.margin=unit(c(0,0,1,0), "cm")) +
    labs(title =paste0(titles[[ifelse( !(i %% 6==0),as.integer(i/6)+1,i/6)]], "- ",year[i]), subtitle = "Cooccurrences within same sentence (Nouns & Adjective)")
}
#show the plots
plotyear <- Filter(Negate(is.null),plotkey)
ggarrange(plotlist=plotyear,nrow=28,ncol=1)
```
# [Quantitative questions](index.html)



### Session Info

```{r}
sessionInfo()
```


