---
title: "BqExAvI Survey Analysis (2017-2022)"
author: "Modesto"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 2
    number-sections: true
    code-overflow: wrap
    code-fold: true
    code-summary: "Show the code"
#output:
#  pdf_document:
#    latex_engine: xelatex
---

# Contents and Disclaimer

This file contains the data from the "Experimental Advanced Biochemistry I" course (Biochemistry Degree, [Universidad Autónoma de Madrid](https://www.uam.es)). The questionnaires were completed by the students in 2017-2022 directly on Moodle on the last day of each course. This is a preliminary summary of the data analysis.  The GitHub [repo](https://github.com/mredrejo/bqexav) contains the original files of all analyzes.


The data is made available under the Creative Common License (CC BY-NC-ND 3.0 ES).  Contact [modesto.redrejo\@uam.es](mailto:modesto.redrejo@uam.es) for more information.

# Moodle survey

```{r include=FALSE,warning=FALSE}
#Load/install requires packages
paquetes <- c("ggplot2","data.table","kableExtra","corrplot","likert","ggpubr","heatmaply","reshape2","plotly","dplyr")
unavailable <- setdiff(paquetes, rownames(installed.packages()))
invisible(install.packages(unavailable))
invisible(lapply(paquetes, library, character.only = TRUE))
```

We import data from student responses in 5 years, between 2018-2022. The quiz consists of up to 77 questions, including 5 free text questions (50, 51, 52, 75, 76 & 77), 7 questions with three options and the rest on a Likert scale. The latter are categorized as "Positive" or "Negative" where 5 ("strongly agree") means "Good" or "Bad" feedback.

```{r warning=FALSE}
#load questions
questions <- read.csv("questions.csv", head=TRUE, sep=";")
questions <- cbind(row.names(questions),questions[,c(3,1,2)])
#add type variable
questions$type <- "Pos."
questions$type[questions$Section=="Open"] <- ""
questions[c(4,10,56,63,63,64,65,66,67,72),5] <- "Neg."
questions$type[questions$Section=="Open"] <- "NA"
colnames(questions) <- c("No.","Since","Section","Question","Type")

kbl(questions, align = "cccl", caption = "Table 1. Students opinion quizz. The bulk of the questionaire was designed for the year 2017 and new questions were added as indicated in the rightmost column.") %>%
    kable_styling(bootstrap_options = "striped", full_width = F) %>%
    column_spec(1, italic = T)
```

# Load data and pairwise t.test

Survey responses were downloaded from Moodle as txt/csv files. Moodle updates caused some format differences that could be worked around after opening the files with Numbers and exporting them as tables with ";" as the column separator. We now show pairwise t-tests performed to detect significant differences in responses per year.

```{r warning=FALSE}
#read the data in a list of dataframes
#didn't use the headers to avoid mistakes
quiz <- lapply(2017:2022, function(x) read.csv(paste0("survey",x,".csv"),header=FALSE,skip=1,sep=";"))
#add Year as the third variable (empty so far)
curso <- c("2017","2018","2019","2020","2021","2022")
for (i in 1:length(quiz)){
  quiz[[i]][,3] <- curso[i]
}

#adjust questions changes
#remove questions in column 32 & 40 from 2017, because we removed it in the following years
quiz[[1]] <- quiz[[1]][,-c(32,40)]
quiz[[6]] <- quiz[[6]][,-10]
names(quiz[[6]]) <- names(quiz[[5]])
names(quiz[[1]]) <- names(quiz[[2]])
#merge all dataframes and name the columns 
data <- Reduce(function(x, y) merge(x, y, all=TRUE), quiz)
#take the colnames from the last quiz that contains all the questions

colnames(data)[3] <- "Curso"


#statistics analysis
#subset questions 1: remove leftmost junk columns
subdata_all <- data[,c(3,11:87)]
names(subdata_all) <- c("Curso",paste0("Q",1:77))
#subset questions 2: remove open questions
open <- c(row.names(questions[questions$Section=="Open",]))
subdata <- subdata_all[,-(as.integer(open)+1)]
subdata <- sapply(subdata,as.numeric)
subdata <- as.data.frame(subdata)


tests <- list()
nombres <- c()

for (i in 2:ncol(subdata)){
  subdata[,i][!(subdata[,i] %in% c(1,2,3,4,5))] <- NA
  #subset for years with answers to avoid void groups
  kkk <- subset(subdata,!is.na(subdata[,i]))
  tests[[i-1]] <- pairwise.t.test(x=as.numeric(kkk[,i]),g=as.numeric(kkk[,1]),paired = F)
  nombres[i-1] <- questions[,4][(which(questions$No. %in% gsub("\\D", "",colnames(subdata[i]))))]
  names(tests) <- nombres
  print(names(tests[i-1]))
  print(tests[[i-1]]$p.value)
}


```

# Boxplots

Responses to each question are presented in boxplots by year, and significant differences from overall responses are marked.

```{r warning=FALSE}


#update question nubers
#stack table and add question type (except open)
subdata.stack <- cbind(subdata[,1],stack(subdata[,2:72]))
names(subdata.stack) <- c("Curso","value","question")
#add categories
for (i in 1:nrow(subdata.stack)){
  subdata.stack[i,4] <- questions$Section[subdata.stack[i,3]==paste0("Q",questions$No.)]
}
names(subdata.stack) <- c("Curso","value","question","section")

pp <- list()
plot <- list()

#plot in groups by section


for (i in 1:length(levels(as.factor(subdata.stack$section)))){
  kkk <- subdata.stack[subdata.stack$section==levels(as.factor(subdata.stack$section))[i],]
  pp[[i]] <- ggboxplot(kkk, x = "Curso", y = "value",
          color = "Curso", fill = "Curso",palette = "jco",
          add = "jitter",alpha = 0.3, ylim = c(0, 6),
          facet.by = "question")
  plot[[i]] <- pp[[i]] + ggtitle(levels(as.factor(subdata.stack$section))[i]) + theme(axis.text.x = element_text(size = 15, angle=45,vjust=0.7),axis.text.y = element_text(size = 15),plot.title=element_text(hjust=0.5,face="italic")) + theme(legend.position="none") +
   stat_compare_means(label = "p.signif",method = "t.test", ref.group = ".all.", hide.ns = TRUE,label.y=5.5,size=6) 
}

```

```{r  echo=TRUE, fig.height=12,warning=FALSE}
plot[[1]]
plot[[2]]
```

```{r  echo=TRUE, fig.height=8,warning=FALSE}
plot[[3]]
```

```{r  echo=TRUE, fig.height=20,warning=FALSE}
plot[[4]]

```

```{r  echo=TRUE, fig.height=4,warning=FALSE}
plot[[5]]
```

```{r  echo=TRUE, fig.height=8,warning=FALSE}
plot[[6]]
```

```{r  echo=TRUE, fig.height=12,warning=FALSE}
plot[[7]]
```

```{r  echo=TRUE, fig.height=4,warning=FALSE}
plot[[8]]
```

```{r  echo=TRUE, fig.height=12,warning=FALSE}
plot[[9]]

```

# Likert scale plots

Likert scale responses are represented by sections in the quiz.

## General Methodology

```{r Fig2, echo=TRUE, fig.height=7,warning=FALSE}

#lickert
#change question names
tablita <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = length(colnames(subdata)),
                          ncol = 2))
for (i in 2:length(colnames(subdata))){
  tablita[i-1,] <- cbind(colnames(subdata[i]),questions$Question[as.numeric(gsub("\\D", "",colnames(subdata[i])))])
  colnames(subdata)[i] <- tablita[i-1,2]
}

for (i in 2:72){
  subdata[,i] <- factor(subdata[,i])
}
subdata$Curso <- factor(subdata$Curso,levels=c(2022,2021,2020,2019,2018,2017))
#questions with 3 options
#General Methodology
subdata[,c(8:12)] <- lapply(subdata[,c(8:12)], function(x) factor(x, 
      labels = c("Traditional","Same","Open Question"))
  )
xlikgroup3a = likert(subdata[,c(8:12)], grouping = subdata$Curso)
plot(xlikgroup3a, type = "bar", centered = T) 
```

## Activities Length

```{r Fig3, echo=TRUE, fig.height=10,warning=FALSE}
subdata[,c(29:35)] <- lapply(subdata[,c(29:35)], function(x) factor(x, 
      labels = c("Less","Fine","More"))
  )

xlikgroup3b = likert(subdata[,c(29:35)], grouping = subdata$Curso)

plot(xlikgroup3b, type = "bar", centered = T)
#title(main = "Activities Length", xlab = "X axis", ylab = "Y axis", cex.main = 4,   font.main = 3)
#legend("bottom",  c("Less","Fine","More"))
```

## Equipment

```{r Fig4, echo=TRUE,warning=FALSE}
subdata[,c(2:4)] <- lapply(subdata[,c(2:4)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5a = likert(subdata[,c(2:4)], grouping = subdata$Curso)
plot(xlikgroup5a, type = "bar", centered = T)

```

## Length and Schedule

```{r Fig5, echo=TRUE, fig.height=5,warning=FALSE}
subdata[,c(5:7)] <- lapply(subdata[,c(5:7)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
xlikgroup5b = likert(subdata[,c(5:7)], grouping = subdata$Curso)
plot(xlikgroup5b, type = "bar", centered = T)

```

## Method Objectives

```{r Fig6, echo=TRUE, fig.height=13,warning=FALSE}
subdata[,c(13:21)] <- lapply(subdata[,c(13:21)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
xlikgroup5c = likert(subdata[,c(13:21)], grouping = subdata$Curso)
plot(xlikgroup5c, type = "bar", centered = T)

```

## Activities Interest

```{r Fig7, echo=TRUE, fig.height=9,warning=FALSE}
subdata[,c(22:28)] <- lapply(subdata[,c(22:28)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5d = likert(subdata[,c(22:28)], grouping = subdata$Curso)
plot(xlikgroup5d, type = "bar", centered = T)

```

## Assessment

```{r Fig8, echo=TRUE, fig.height=9,warning=FALSE}
subdata[,c(36:41)] <- lapply(subdata[,c(36:41)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5e = likert(subdata[,c(36:41)], grouping = subdata$Curso)
plot(xlikgroup5e, type = "bar", centered = T)

```

## Learning Objectives

```{r Fig9, echo=TRUE, fig.height=12,warning=FALSE}
subdata[,c(43:50)] <- lapply(subdata[,c(42:50)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )

xlikgroup5f = likert(subdata[,c(42:50)], grouping = subdata$Curso)
plot(xlikgroup5f, type = "bar", centered = T)

```

## ELN

```{r Fig10, echo=TRUE, fig.height=28,warning=FALSE}
#subset to remove empty years
subdata[,c(51:72)] <- lapply(subdata[,c(51:72)], function(x) factor(x, 
      labels = c("Not at all","Disagree","OK","Agree","Completely Agree"))
  )
eln <- subset(subdata[,c(1,51:72)][subdata$Curso==2019|subdata$Curso==2020|subdata$Curso==2021|subdata$Curso==2022,])


xlikgroup5g = likert(eln[2:23], grouping = eln$Curso)
plot(xlikgroup5g, type = "bar", centered = T, title="ELN")
```

### Session Info

```{r}
sessionInfo()
```
